---
title: "Time Series"
author: "Léonard Gousset"
date: "2023-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Libraries importation

```{r}
library(readr)
library(readxl)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(forecast)
library(TSA)
library(caschrono)
library(tseries)
library(fracdiff)
```

## 1) Data Preprocessing: Clean and preprocess the data. This may involve handling missing values, converting timestamps, and normalizing or scaling the data.

The dataset under consideration, known as the Sea Ice Index, is a valuable collection providing a focused examination of sea ice dynamics. Comprising two key variables—namely, "Date" and "Extent"—the dataset allows for a comprehensive temporal analysis of sea ice behavior. The "Date" variable encapsulates temporal information, enabling the exploration of sea ice patterns over distinct time points. Concurrently, the pivotal "Extent" variable quantifies the sea ice coverage, measured in square kilometers. This metric serves as a crucial indicator, offering insights into the spatial dynamics of sea ice. Through the utilization of the Sea Ice Index dataset, this project aims to uncover trends, seasonality, and anomalies in sea ice extent. The analysis will not only enhance our understanding of the intricate variations in sea ice over time but also contribute to the broader discourse on climate change impacts, specifically in relation to diminishing sea ice coverage. 

```{r, echo=FALSE}
# Dataset importation
raw_db <- read_xlsx("N_ice.xlsx")
raw_db_copy <- raw_db
```

```{r}
# Transform the month, day, year to one feature : Date
db <- raw_db %>% 
  mutate(Date = as.Date(paste(Year, Month, Day, sep = "/"))) %>% 
  select(Extent, Date, Year, Month)
db$Extent <- as.numeric(db$Extent)
summary(db)
```
```{r}
sum(is.na(db))
```
We don't have any missing values so we can start our Exploratory Data Analysis.

## 2) Exploratory Data Analysis (EDA): Explore the data to understand its characteristics, identify trends, seasonality, and any potential anomalies or outliers.

```{r}
# transformation of the feature "Extent" to a time series object
extent_ts <- ts(db$Extent, frequency = 1, start = c(db$Year, db$Month))

# See the time series
ggplot(data = db, aes(x = Date, y = Extent)) +
  geom_line() +
  labs(x = "Date", y = "Extent") +
  ggtitle("time Series of Extent") +
  theme_bw()
```
On this graph, we can see two major things. The first one is seasonality, 

```{r}
# Plotting the distribution of Sea Ice Extent
ggplot(data = db, aes(x = Extent)) +
  geom_histogram(fill = "#6A7073", color = "black", alpha = 0.7, bins = 50) +
  labs(title = "Distribution of Sea Ice Extent",
       x = "Sea Ice Extent (Square Kilometers)",
       y = "Frequency") +
  theme_bw()

```

